---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

## Loading and preprocessing the data
### Load library 
```{r, message=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
library(lattice)
```

### Load/check data
```{r, message=FALSE}
FileName <- 'activity.csv'
df <- read.table(FileName, sep = ',', header = TRUE, quote = "\"", na.strings = "", stringsAsFactors = FALSE, colClasses = c("numeric", "character","integer"))
str(df)

dfsub1 <- subset(df, !is.na(df$steps)) # subset of orginal data without missing value
str(dfsub1)
```

## What is mean total number of steps taken per day?
First, calculate the total number of steps taken per day, after ignore the missing value:
```{r, message=FALSE}
stepsPerDay <- summarize(group_by(dfsub1, date), stepsPerDay = sum(steps))
print(stepsPerDay)
```

Histogram of the total number of steps taken per day (used default bin size and width):
```{r, message=FALSE}
ggplot(stepsPerDay, aes(stepsPerDay)) + 
        geom_histogram() +
        labs( x = "Total Steps per day") +
        labs( y = "Frequency")
```

The mean and median are:
```{r, message=FALSE}
mean(stepsPerDay$stepsPerDay)

median(stepsPerDay$stepsPerDay)
```
## What is the average daily activity pattern?
calculate the average number of steps taken across all days and then plot in a times series
```{r, message=FALSE}
stepsInt <- summarize(group_by(dfsub1, interval), stepsInt = mean(steps))
print(stepsInt)
plot(stepsInt$steps ~ stepsInt$interval, type="l", xlab="Interval(5-mins)", ylab="Average Steps across days")
```

find the row of 5-minute interval that contains the maximum number of steps
```{r, message=FALSE}
stepsInt[which(stepsInt$steps == max(stepsInt$steps)), ]
```

## Imputing missing values
### Calculate the total number of missing value in the original dataset
```{r, message=FALSE}
sum(is.na(df)) 
```
### Imputate all the missing values in the dataset, with the mean value in that interval across days
```{r, message=FALSE}
dfNew <- df 
for (i in 1:nrow(dfNew)) {
    if (is.na(dfNew$steps[i])) {
        dfNew$steps[i] <- stepsInt$stepsInt[which(dfNew$interval[i] == stepsInt$interval)]
    }
}
sum(is.na(dfNew))
str(df) # old dataset summary
str(dfNew) # new dataset summary
```

### Histogram of total number of steps/day after imputation
```{r, message=FALSE}
stepsPerDayNew <- summarize(group_by(dfNew, date), stepsPerDayNew = sum(steps))
print(stepsPerDayNew)
ggplot(stepsPerDayNew, aes(stepsPerDayNew)) + 
        geom_histogram() +
        labs( x = "Total Steps per day after imputation") +
        labs( y = "Frequency") +
        labs(title = "New dataset")
ggplot(stepsPerDay, aes(stepsPerDay)) + 
        geom_histogram() +
        labs( x = "Total Steps per day") +
        labs( y = "Frequency") +
        labs(title = "Old dataset")

```

## Are there differences in activity patterns between weekdays and weekends?

```{r, message=FALSE}
dfNew <- mutate(dfNew, day = factor(weekdays(as.Date(dfNew$date))))
dfNew <- mutate(dfNew, weekday = factor(ifelse(dfNew$day %in% c("Saturday", "Sunday"), "Weekend", "Weekday")))
head(dfNew)
str(dfNew)

stepsIntWeekday <- summarize(group_by(dfNew, interval, weekday), stepsIntWeekday = mean(steps))
head(stepsIntWeekday)
str(stepsIntWeekday)

xyplot(stepsIntWeekday ~ interval | weekday, stepsIntWeekday, type = "l", layout = c(1, 2), xlab="Interval(5-mins)", ylab="Average Steps")
```

